Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-09 edubots_surveys.Rnw
incorporates changes from previous review PR#84

Given the potential benefits that chatbots could bring to the classroom, the
focus of the % EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
unnamed European project, the main sponsor of this study, is to explore best practices and innovative use of chatbots,
and to create a learning community of educators in higher education institutions.
Researchers and teachers participating in this project have developed several studies. Most of these works have been focused on how the
messaging applications and/or chatbots are normally applied to deliver personalized learning in classrooms that occurs anytime anywhere,
promote collaborative learning experiences, group discussions \cite{panah2020study} and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}.
The majority of these studies were conducted focusing on the teachers' perspectives, so this paper aims to provide more conclusions regarding
their opinions and needs, as well as the existing challenges or opportunities that arise when adopting these technologies in classrooms. This will lead to clarify
how they can positively contribute to enhancing the learning process in higher education institutions.

@@

WhatsApp is, at least in most Western countries, the most popular MIM platform
used by educators to provide assignments' feedback to students, support course
discussions, and contribute resources in informal learning settings
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

>>>>

Given the potential benefits that chatbots could bring to the classroom, the
focus of the % EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
unnamed European project, the main sponsor of this study, is to explore best practices and innovative use of chatbots,
and to create a learning community of educators in higher education institutions.
Researchers and teachers participating in this project have developed several studies. Most of these works have been focused on how 
messaging applications and/or chatbots are normally applied to deliver personalized learning in classrooms that occurs anytime anywhere, and
promote collaborative learning experiences or group discussions \cite{panah2020study}, and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}.
The majority of these studies were conducted focusing on the teachers' perspectives, so this paper aims to provide more conclusions regarding
their opinions and needs, as well as the existing challenges or opportunities that arise when adopting these technologies in classrooms. This will lead to clarify
how they can positively contribute to enhancing the learning process in higher education institutions.

@@

WhatsApp is, at least in most Western countries, the most popular MIM platform. The reason is that it is the tool preferably
used by educators to provide assignments' feedback to students, support course
discussions, and post learning resources in a semi-formal learning context
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education, as well as applications based on them such as chatbots, is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell in
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
using chatbots to collect course feedback from students in higher
education improved the quality of responses given by students in their assignments, and boosted engagement levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.
